Verifying the \shaThree\ Implementation from \openssl\ with the Software Analysis Workbench

### 2022-01-08 paper/openssl.tex
Fix figures

\begin{figure}
  \begin{center}
    \usebox{\PiOpenSSL}
  \end{center}
  \caption{The \openssl\ implementation of $\pi$ that reverses indexes and operates on lanes.}
  \label{fig:piopenssl}
\end{figure}

@@

\newsavebox{\cryciota}
\begin{lrbox}{\cryciota}
  \begin{lstlisting}[language=Cryptol]
    LISTIOTAS = [reverse (
      (while {RC = 0:[W], j = 0} 
              (\state -> state.j <= `L) 
              (\state -> { 
                RC = for [0..63] 
                  (\z -> if z == index then rc (state.j + 7*i)
                          else state.RC @z)
                    where index = ((1:[8]) << state.j) - 1,
                                    j = state.j + 1})
      ).RC) | i <- LISTROUNDS:[_][64]]

    LISTIOTAS = [
      0x0000000000000001, 0x0000000000008082, 0x800000000000808a,
      0x8000000080008000, 0x000000000000808b, 0x0000000080000001,
      0x8000000080008081, 0x8000000000008009, 0x000000000000008a,
      0x0000000000000088, 0x0000000080008009, 0x000000008000000a,
      0x000000008000808b, 0x800000000000008b, 0x8000000000008089,
      0x8000000000008003, 0x8000000000008002, 0x8000000000000080,
      0x000000000000800a, 0x800000008000000a, 0x8000000080008081,
      0x8000000000008080, 0x0000000080000001, 0x8000000080008008]
        :[_][64]
  \end{lstlisting}    
\end{lrbox}

\begin{figure}[t]
  \begin{center}
    \begin{tabular}{l}
      \usebox{\cryciota} \\ \\
    \end{tabular}
  \end{center}
  \caption{Optimized $\iota$ function. (a) \openssl's implementation. (b) Matching \cryptol\ implementation.}
  \label{fig:iota}
\end{figure}

The second difference is in hardcoding lookup Tables for constants rather than computing the constants along the way.
The \fips\ \cryptol\ specification computes each constant on-the-fly as needed in each round.
This redundant computation is seen with the rounding constants, \emph{rc}, for $\iota$, and the rotate constants for $\rho$. 
The \openssl\ implementation hardcodes these constants in tables without the provenance of the computation and looks up the constants in the tables.
The new \cryptol\ specification for the \openssl\ uses the same lookup tables, but it differs from the \openssl\ implementation in that it computes the lookup tables once so that the values in the table now have provenance.
This provenance is sufficient for the equivalence proof with the C code.
\figref{fig:iota} shows the difference in the two specifications for the rounding constants with the code for the computed table coming directly from \fips\ in the \openssl\ specification.

>>>>

\begin{figure}[t]
  \begin{center}
    \usebox{\PiOpenSSL}
  \end{center}
  \caption{The \openssl\ implementation of $\pi$ that reverses indexes.}
  \label{fig:piopenssl}
\end{figure}

@@

\newsavebox{\cryciota}
\begin{lrbox}{\cryciota}
  \begin{lstlisting}[language=Cryptol]
    iota a i =  for LIST4 (\y ->
                  for LIST4 (\x ->
                    for LIST63 (\z ->
                      if ((x == 0) && (y == 0))
                        then (a @0 @0 @z) ^ (LISTIOTAS @i @z)
	                      else a @y @x @z)))

    LISTIOTAS = [reverse (
      (while {RC = 0:[W], j = 0} 
              (\state -> state.j <= `L) 
              (\state -> { 
                RC = for [0..63] 
                  (\z -> if z == index then rc (state.j + 7*i)
                          else state.RC @z)
                    where index = ((1:[8]) << state.j) - 1,
                                    j = state.j + 1})
      ).RC) | i <- LISTROUNDS:[_][64]]
  \end{lstlisting}    
\end{lrbox}

\begin{figure}[t]
  \begin{center}
    \begin{tabular}{l}
      \usebox{\cryciota} \\ \\
    \end{tabular}
  \end{center}
  \caption{The computed \emph{rc} table with the lookup in $\iota$.}
  \label{fig:iota}
\end{figure}

The second difference is in hardcoding lookup Tables for constants rather than computing the constants along the way.
The \fips\ \cryptol\ specification computes each constant on-the-fly as needed in each round.
This redundant computation is seen with the rounding constants, \emph{rc}, for $\iota$, and the rotate constants for $\rho$. 
The \openssl\ implementation hardcodes these constants in tables without the provenance of the computation and looks up the constants in the tables.
The new \cryptol\ specification for the \openssl\ uses the same lookup tables, but it differs from the \openssl\ implementation in that it computes the lookup tables once so that the values in the table now have provenance.
This provenance is sufficient for the equivalence proof with the C code.
\figref{fig:iota} shows the $\iota$ function that uses the computed lookup table.
